
## Title: "Bicycle Theft data analysis"
## Author: Suyash Gautam"
```{r}
library(lattice)
library(ggplot2)
library(ggmap)
```

##FILE IMPORT  

(I split the 'geom' varible in the excel file before import to make spatial plotting easy)

```{r}
btheft <- read.csv("~/Documents/R_dir/Bicycle_Thefts_9_15_2014_-_2_8_2015 (1).csv")
head(btheft)
```

# Working with date
# The data under observation is from September 15, 2014 to February 8, 2015

```{r}
class(btheft$THEFT_DATE)
new_date <- as.Date(btheft$THEFT_DATE, format = '%m/%d/%Y')
btheft$THEFT_DATE <- new_date
class(new_date)
head(new_date)
summary(new_date)
```


# Working with Months 

```{r}
month <- months(new_date)
countmonth <- table(month)
countmonth <- sort(countmonth, decreasing = FALSE)
countmonth
prop.table(countmonth)
barplot(countmonth, main="Theft distribution by month", 
        xlab = "Month", pch = 18, ylab = "Thefts", col = "red", lwd = 2)
```

reported thefts in October -> 320, followed by 
September -> 209 
November -> 127 thefts.  
No thefts reported in March, April, May, June, July and August. 

# Working with time

```{r}
month_number <- as.numeric(format(new_date, "%m"))
qplot(month_number, btheft$THEFT_HOUR, xlab = "Month", ylab = "Theft Hour",
      main = "Theft hour by Month", col = THEFT_HOUR, data = btheft)
```

# Now lets study distribution of theft hour

```{r}
hist(btheft$THEFT_HOUR, xlab = "Hour", ylab = "Frequency", col = "blue", 
     main = "Histogram of Theft Hour", lwd = 2)
lines(density(btheft$THEFT_HOUR), lwd = 2, col = "chocolate")
abline(v = median(btheft$THEFT_HOUR), col = "red", lwd = 2)
hourtable <- table(btheft$THEFT_HOUR)
#View(hourtable)
ht <- prop.table(hourtable)
#View(ht)
summary(btheft$THEFT_HOUR)
boxplot(btheft$THEFT_HOUR, col = 'maroon', main = "Boxplot of theft hour")
```


50% of the thefts reported to have occured before 4:00 pm.  
75% of the thefts occured between 12:00 AM and 6:00 PM.


 


# Theft hour vs stolen Value
```{r}
boxplot(btheft$STOLEN_VAL~btheft$THEFT_HOUR, data = btheft, xlab = "Reported Time of Theft", ylab = "Stolen Value($)",
      main = "Stolen Value vs Theft time", col = "red")
modelvh <- lm(btheft$STOLEN_VAL ~ btheft$THEFT_HOUR, data = btheft)
abline(modelvh, median(btheft$STOLEN_VAL))
summary(modelvh)
```
A very high p value indicates practically no causal relationship between theft hour and stole value

# Lets subset the data to calculate stolen amounts with respect to hour of the day
```{r}
sum(btheft$STOLEN_VAL)
timed <- subset(btheft, THEFT_HOUR < 12 )
sum(timed$STOLEN_VAL)
timee <- subset(btheft, THEFT_HOUR >= 12 & THEFT_HOUR <= 18)
sum(timee$STOLEN_VAL)
```

All the thefts involving value greater than $1000 occured between 8:00 am and 11:00 pm. 
Total Value reported to be stolen during the period was $347538. 
Total Stolen value of thefts reported to have occured before noon was $83644 or 24% of the 
total stolen amount. Total Stolen value of thefts reported to have occured between noon and 6:00 PM was $177082 or 51% of the total stolen amount.


# Working with days of the month
```{r}
day <- weekdays(new_date) 
countday <- table(day)
countday
prop.table(countday)
barplot(countday, main="Theft distribution by weekday", 
        xlab = "Weekday", ylab = "Thefts", col = "green")
```




We note from the frequency counts as well as from the barplot that thefts are almost evenly 
distributed during the week for the period under study


Now lets try to relate month_number and Stolen_val to see if there is any relationship
```{r}
boxplot(btheft$STOLEN_VAL~month_number, main = 'Stolen Value by month',
        xlab = 'Month', ylab = 'Stolen Value ($)', col = 'red', lwd = 1.5, pch = 9)
```




It seems that months with greater number of thefts also have larger stolen values.




```{r}
with(btheft, plot(month_number, STOLEN_VAL, xlab = "Month", ylab = "Stoeln Value in $",
                  main = 'Stolen Value vs Month'))
modelx <- lm(STOLEN_VAL ~ month_number, data = btheft)
abline(modelx, lwd = 2)
summary(modelx)
```
Again, The pretty straight abline and a 'p' value much larger than 0.05 indicate practically 
no causality here between the month and stolen value. 


# Some more calculations

```{r}
median_by_month <- aggregate(STOLEN_VAL ~ month_number, btheft, median) 
median_by_month
```
The median stolen value by month was close to $300

```{r}
wekd_nb <- as.POSIXlt(new_date)$wday
head(wekd_nb)
table(wekd_nb)
table(day)
daytable <- table(day) 
#View(daytable)
prop.table(daytable)
with(btheft, plot(wekd_nb, STOLEN_VAL, xlab = "weekday", ylab = "stolen Value in $",
                  main = 'Stolen Value vs week day'))
modelz <- lm(STOLEN_VAL~wekd_nb, data = btheft)
abline(modelz, lwd = 2)
summary(modelz)

```
Again, straight abline and a 'p' value of 0.3337 indicate practically 
no causality here between the weekday and stolen value. 

```{r}
median_by_wekd_nb <- aggregate(STOLEN_VAL ~ wekd_nb, btheft, median)
median_by_wekd_nb
```
The median stolen value by weekday was also close to $300


# Working with the distance variable
```{r}
hist(btheft$DC_DIST, xlab = "Distance of theft from the victim", ylab = "Frequency", col = "violet",                   main = "Histogram of distance", lwd = 2)
lines(density(btheft$DC_DIST), lwd = 2, col = "chocolate")
abline(v = median(btheft$DC_DIST), col = "red", lwd = 2)
disttable <- table(btheft$DC_DIST)
#View(disttable)
dt <- prop.table(disttable)
#View(dt)
summary(btheft$DC_DIST)
boxplot(btheft$DC_DIST, col = 'maroon', main = "Boxplot of distance")
```

We note from the frequency counts that 50% of the reported thefts occured less than 16 miles from the victim. 75% of the reported thefts occured between 6 and 18 miles from the victim. 




# Spatial plotting part 1

```{r}

btheft$lat <- gsub("\\(", "", btheft$lat)
btheft$long <- gsub("\\)", "", btheft$long)
class(btheft$lat)
class(btheft$long)

btheft$lat <- as.numeric(btheft$lat)
btheft$long <- as.numeric(btheft$long)
library(ggplot2)
library(ggmap)
philyMap <- qmap("philadelphia", zoom = 15, legend = "topleft", maptype = "roadmap")
plotd <- btheft[, 11:12]
#class(plotd)
#class(plotd$lat)

#View(plotd)
philyMap + geom_point(aes(x = btheft$lat, y = btheft$long), data = btheft)
```



# Spatial Plotting part 2
(I spent a lot of time on this part. This code seems to be causing problems on may mac. I hope it works on your system)

```{r}
#install.packages("ggmap")
library(ggmap)
#install.packages("googleVis")
library(googleVis)
#View(btheft)
placeNames <- as.character(btheft$LOCATION_B) 
head(placeNames)
plotData <- data.frame(name = placeNames, lat = btheft$lat,  long = btheft$long)
head(plotData)

#View(plotData)
#class(plotData)
sites <- gvisMap(plotData, locationvar = "lat", "long", tipvar = "name",
                 options = list(displayMode = "Markers", mapType = "terrain", 
                                colorAxis = "{colors:['red', 'blue']}", height = 1000,
                                useMapTypeControl=TRUE, enableScrollWheel='TRUE'))       
plot(sites)
```

# Insights from the data

 1) A total of 804 thefts were reported for the period from September 15, 2014 to February 8, 2015.
    October had the most number of reported thefts, 320, followed by September with 209 thefts and      November with 127 thefts. February observed the least reported activity. So, October was the        most active theft month with 40% of the thefts reported. We also note that there were no thefts     reported in March, April, May, June, July and August.  

 
 2) It was found that 50% of the thefts were reported to have occured before 4:00 pm. 75% of the        thefts occured between 12:00 AM and 6:00 PM. 

 
 3) A test was conducted to study wheter the stolen value reported increases in later hours of the      day. No causal relationship was found between theft hour and stole value.

 
 4) All the thefts involving value greater than $1000 were found to have occured between 8:00 am        and 11:00 pm. The total $ Value reported to have been stolen during the period under study was      $347538. Total Stolen value of thefts reported to have occured before 12:00 pm was $83644 or        24% of the total stolen amount. Total Stolen value of thefts reported to have occured between       12:00 PM and 6:00 PM was $177082 or 51% of the total stolen amount.

 
 5) While studing the distribution of thefts during the week, tt was found that thefts were almost      evenly distributed during the week for the period under study

 
 6) While studying how thefts related to distance from the victim, it was observed that 50% of the      reported thefts occured less than 16 miles from the victim. 75% of the reported thefts occured      between 6 and 18 miles from the victim.
 
 
 
 
 

